The human in the loop: volunteer computing as a socio-technical system

### 2016-02-28 Alife16/paper.tex
Taking out verbose explanations
Managed to reduce it to 8 pages
Closes #99 but we'll have to keep it that way.

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\citep{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset. This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript.

As a volunteer based system, there are some issues worth 
noting \citep{sarmenta2001volunteer} :\begin{enumerate}
\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
each request is known.
\item As anonymous entities, volunteers are not accountable. 
If a volunteer misbehaves in some way the provider cannot 
prosecute or discipline the volunteer. In the case of JavaScript,
this issue is aggravated because even if it is obfuscated, source code
and data is sent to clients. 
\item Participants must also trust the application provider, 
regarding invasion of privacy, the integrity of their devices, 
and intended application. 
\end{enumerate}

Two kind of threats are possible in this environment; first, since the
API is open and it is an open source framework, it is relatively easy to
sabotage the system, as indicated in \citep{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome.
% (Paloma) Where is the second kind of threat?
Another two different attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released. 
 We also release as open access the performed experiments \citep{DBLP:journals/corr/GuervosG15} 
 and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade the
performance of the system. This socio-technical approach is, indeed,
coherent with our whole approach to the study of a distributed volunteer
 system which can only be studied by approaching it from the
technical as well as the social, as in social media, angle.

@@

In this case the classical Trap function \citep{Ackley1987} has been
used. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic}, is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):
\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Node.js and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a non-blocking single thread
  allows the service of many requests. In fact, a limit in the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \citep{gecco07:workshop:dcor}, which had a low scalability. 
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system, or even browser requirement as long
  as JavaScript is enabled: anyone
  visiting the page, even from mobile devices, can load the algorithm.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through the autonomous
  operation of every individual island without any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. This architecture 
  has been tested in other high demand systems with success, but 
  in the case of {\sf NodIO}, additional experiments are needed to asses 
  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
  %(Mario) Scalability of the communication system in particular?
\item Since the algorithm runs in the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources than the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which makes the operation totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL, available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

There are many different ways to validate a framework that intends to
address these challenges. Some of them are included by design: it is
anonymous, since no accounts are needed to participate in the
experiment, it is safe for the user, since we are using the browser
black box, and it is also heterogeneous. We have designed the
experiments so that they show adaptability to different types of
experiments and clients, scalability with the number of clients and
the problem size, and measured performance by comparing it with
a baseline configuration, with which, as indicated above, is
compatible and can be used in a complementary way. Fault tolerance
will be experimented by dropping the server and checking the continuity
of the experiments.

All these points will be established in the next two sections,
starting with the baseline, single worker system next. %%%  Single worker?

>>>>

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. We should emphasize
that NodIO is more intended as an auxiliary computing engine, more
than the main one, so performance of JavaScript as a language is not
so important; even so, we have made a comparison between JavaScript and other languages
\citep{2015arXiv151101088M} that shows that performance of JavaScript
is comparable to other interpreted languages; compiled languages would
be faster, but, of course, it is impossible to gather volunteers
spontaneously and without any installation with them.

@@

In this case the classical Trap function \citep{Ackley1987} has been
used. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic}, is the only change needed to work with a different
problem. The next Section will describe experiments performed to
establish baseline performance and gather initial performance
results.
